---
title: "Project 1 week 2"
author: "Gary"
date: "28/09/2020"
output: html_document
---
## Set directory and import libraries
```{r}
setwd("C:/Users/Gary Clarke/Desktop/Coursera")
library(ggplot2)
library(readr)
library(dplyr)
```

## Import file as a dataframe
```{r}
activityDT <- data.table::fread(input = "activity.csv")
```
## Calculate and plot the histogram to show the total number of steps taken each day
```{r}
Total_steps <- activityDT[, c(lapply(.SD, sum, na.rm = FALSE)),
                          .SDcols = c("steps"),
                          by = .(date)]

ggplot(Total_steps, aes(x=steps))+
               geom_histogram(fill = "green", binwidth = 1000)+
               labs(title = "Daily steps", x = "Steps", y = "Frequency")+
               theme_classic()
```


## Calculate the mean and median steps taken each day
```{r}
Total_steps[, .(Mean_steps = mean(steps, na.rm = TRUE), 
                Median_steps = median(steps, na.rm = TRUE))]
```
## Calculate and plot a time series of the average number of steps taken per day



```{r}
IntervalDT <- activityDT[, c(lapply(.SD, mean, na.rm = TRUE)),
                         .SDcols = c("steps"), by = .(interval)]

ggplot(IntervalDT, aes(x = interval, y = steps)) + 
        geom_line(color = "red", size =1) + labs(title = "avg daily steps" , 
                x = "interval", y = "avg steps per day" ) +
                        theme_classic()
```






## Identify the 5 minute interval that contains the maximum number of steps(on avg).

```{r}
IntervalDT[steps == max(steps), .(max_interval = interval)]
```
## Identify number of missing values
```{r}
activityDT[is.na(steps), .N]
```
# Fill in missing values with a median of the dataset
```{r}
activityDT[is.na(steps), "steps"] <- activityDT[, c(lapply(.SD, median, 
                                na.rm =TRUE)), .SDcols = c("steps")]
```
## Create new dataset with the missing values included




```{r}

data.table::fwrite(x = activityDT, file = "data/UpdatedData.csv", quote = FALSE)
```
## Plot Histogram of total steps 
```{r}
Total_steps <- activityDT[, c(lapply(.SD, sum)), .SDcols = c("steps"), 
                          by = .(date)]

Total_steps[, .(Mean_steps = mean(steps), Median_steps = median(steps))]

ggplot(Total_steps, aes(x = steps)) +geom_histogram(fill = "red", 
        binwidth = 1000) + labs(title = "Daily Steps", x = "Steps", 
                y = "Frequency")+ theme_classic()
```










# Identify weekdays from weekends and review output
```{r}
activityDT <- data.table::fread(input = "activity.csv")
activityDT[, date := as.POSIXct(date, format = "%Y-%m-%d")]
activityDT[, `Day of Week`:= weekdays(x = date)]
activityDT[grepl(pattern = "Monday|Tuesday|Wednesday|Thursday|Friday", 
                 x = `Day of Week`), "weekday or weekend"] <- "weekday"
activityDT[grepl(pattern = "Saturday|Sunday", x = `Day of Week`),
           "weekday or weekend"] <- "weekend"
activityDT[, `weekday or weekend` := as.factor(`weekday or weekend`)]
head(activityDT, 10)
```
## Plot of Weekdays and Weekends
```{r}
activityDT[is.na(steps), "steps"] <- activityDT[, c(lapply(.SD, median, 
                                na.rm = TRUE)), .SDcols = c("steps")]
IntervalDT <- activityDT[, c(lapply(.SD, mean, na.rm = TRUE)), 
        .SDcols = c("steps"), by = .(interval, `weekday or weekend`)] 

ggplot(IntervalDT , aes(x = interval , y = steps, color=`weekday or weekend`)) + 
        geom_line() + labs(title = "Avg. Daily Steps by Weektype", 
        x = "Interval", y = "No. of Steps") + 
        facet_wrap(~`weekday or weekend` , ncol = 1, nrow=2) +
        theme_classic()

```

